---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r computeTime, echo = FALSE}
time <- format(Sys.time(),"%d %b %Y")
```

- Author : J. Ivan Avalos
- date : `r time`

## Loading and preprocessing the data
* Check if the zip file has been unzipped
* Read activity.csv file
```{r processingData, cache=TRUE}
if(!file.exists("activity.csv")){
    print("Unzipping the file")
    unzip(zipfile = "activity.zip")}

DFactivity <- read.csv("activity.csv",header = TRUE,na.strings = c("NA"))
head(DFactivity)
```


## What is mean total number of steps taken per day?
* Use aggregate function to compute total steps by day
* Histogram of total steps
* Use aggregate function to compute mean and median steps by day
* Compute mean and median of the total number of steps per day
```{r noSteps}
no.steps <- aggregate(steps ~ date, DFactivity, sum, na.rm = TRUE)
```
```{r plots1}
hist(no.steps$steps, xlab = "Total number of steps per day", ylab = "Days",
     main = "Histogram of steps taken each day", col = "green",
     breaks = seq(0,25000,2000), ylim = c(0,20), xlim = c(0,25000))
```

Mean and median number of steps taken each day
```{r MeanMedia}
mean_Steps <- mean(no.steps$steps)
mean_Steps
median_Steps <- median(no.steps$steps)
median_Steps
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
